Title

### 2016-05-22 paper/sections/Discussion.tex
More text Discussion.

- Transformations and Binary methods (but argue that 
these types of operations are not used in shallow DSLs 
any way!).

- Java syntax not great for DSLs. 
operator overloding; infix operators (but this is a Java limitation; 
not a limitation of our approach).

- Limitations of the tool: separate compiltions and ... 
generics not fully implemented.

- How to apply our approach to other OO language 
  - C\# has annotations, but can you do AST rewriting?
  - Scala has macros;
  - Other languages?

>>>>

%Haoyuan writes this one.

%- Transformations and Binary methods (but argue that
%these types of operations are not used in shallow DSLs
%any way!).

%- Java syntax not great for DSLs.
%operator overloding; infix operators (but this is a Java limitation;
%not a limitation of our approach).

%- Limitations of the tool: separate compiltions and ...
%generics not fully implemented.

%- How to apply our approach to other OO language
%  - C\# has annotations, but can you do AST rewriting?
%  - Scala has macros;
%  - Other languages?

\paragraph{Transformations} Our approach tends to give a taste that shallow embeddings are convenient to
use in object-oriented languages, and extensible with family polymorphism. Nevertheless, shallow embeddings are also under
criticism in terms of its not supporting ASTs explicitly, leading to the fact that some operations like transformations
cannot be easily used with shallow embeddings. In Section ?(2.3) we give an example of desugaring in Java using shallow
embeddings, however when we try to apply family polymorphism for inheritance and extensibility, code duplication can be introduced.
Suppose we firstly integrate the four members \textsf{Circuit}, \textsf{Fan}, \textsf{Beside} and \textsf{Identity} in a base family,
and then define a new family that extends the base one, then inside the new member \textsf{Identity}, how can we implement the transformation \textsf{desugar()}? Certainly we cannot use \textsf{super} to call the old implementation, otherwise we are invoking old constructors of \textsf{Beside},
leading to type errors afterwards. For sure we can just copy and paste the code for the old implementation, but too much
code duplication would depress the users. Another solution is to use F-bounded polymorphism (\textcolor{red}{Haoyuan: refs. Need example?
We do have an example in the Pretty Printer case study}),
in that case code duplication is avoided, but accordingly, it arouses explosively heavy uses of types and parameterisations.

\paragraph{Binary methods} Binary methods are indeed another sort of operations that prevent users from using shallow embeddings in Java
comfortably. \textcolor{red}{Haoyuan: refs?} Java supports covariant return types, hence we can refine the types of field methods and 
simple interpretations, but with binary methods like \textsf{equals}, method subtyping in inheritance is a big problem since the argument type
is changed. In Scala, however, we can deal with binary methods with the help of ? \textcolor{red}{Haoyuan: Could you help Weixin?}.

\paragraph{Shallow or deep embeddings?} We know the fact that our approach integrates shallow embeddings and simple family polymorphism, and that in some object-oriented languages like Java,
some trouble is introduced when dealing with extensible transformations or binary methods. But on the other hand, these types of operations are not used in
shallow DSLs anyway; with a single interpretation we use shallow embeddings, but since shallow embeddings do not represent abstract syntax trees,
it is really hard to realize transformations, and similarly, with binary methods, users tend to use deep embeddings instead to build data hierarchies.

\paragraph{Limitations} Our tool has its certain limitations at a few aspects. As Lombok only provides experimental support
for separate compilation, our \textsf{@Family} does not support separate compilation at this stage, so all the related
interfaces should be included in a single Java file. On generics, our implementation of \textsf{@Family} provides support
for generic interfaces and methods without bounds, on the other hand, generic method typing is not explicitly captured
by the annotation but delegated to the Java compiler. \textcolor{red}{Haoyuan: We still need to talk about Lombok in detail?
Another limitation is that only Eclipse handler is supported.}

Regarding syntax, it would be nice to have operator overloading and infix operators in Java, so that the code could be
written in a more concise and elegant way. Nevertheless, this is limited by the programming language, and we expect our approach to be oriented to
other OO languages. \textcolor{red}{Haoyuan: How to apply our approach to other OO language? C\# has annotations, but can you do AST rewriting? Scala has macros;  Other languages?}

### 2016-05-22 paper/sections/Tool.tex
More text Discussion.

In this section, we present an overview of the code that \lstinline{@Family} generates. The syntax and type system are
consistent with the Java language. We use translation functions to illustrate our code generation.
To make it clearer, we split the process of code generation into three parts, in which case we introduce two
new annotations \lstinline{@Obj} and \lstinline{@ObjOf}, which are not implemented in the source code but just help to explain
the process of translation. In our implementation, the annotation processing is a combination of the three parts, corresponding to
\lstinline{@Family}, \lstinline{@Obj} and \lstinline{@ObjOf} respectively. The translation rules are presented in Figure \ref{fig:trans1}
and Figure \ref{fig:trans2}, together with some auxiliary functions.

\subsection{\textsf{@Family}}
\lstinline{@Family} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Family} tackles two
tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
aim, \lstinline{@Family} re-declares all member types and field methods from the base families (see \textsf{collectMembers} and \textsf{fieldMethods} in Figure \ref{fig:trans1}). \lstinline{@Family} recognizes field methods by checking if they are non-void, no-argument methods, and if the name starts with an underscore ``\_''.  The function \textsf{collectMembers} finds all the direct member types from base families $I_1,\cdots,I_n$, creates new types with the same names for re-declaration, and builds the subtyping relations among them. The newly created types are also annotated with \lstinline{@Family}, leading to such a recursive process throughout the nested interfaces in base families. If users have already declared these member types in order to introduce new operations, behaviors or whatever, the function \textsf{combine} will integrate them to the re-declaration, by combining methods and supertypes following ``extends''.

@@

In Figure \ref{fig:trans2}, the first translation function presents the processing of \lstinline{@Obj}. Besides delegating \lstinline{@Obj} to the member types in a recursive way, it invokes the function \textsf{refine} to handle the type refinements. The definition of \textsf{refine} is shown in (1) and (2), where return types of \textsf{with-} methods and fluent setters are refined respectively. They are both state operations, and for simplicity we omit the case for general \textsf{with} method in our translation rules.

@@

\begin{figure*}
\begin{lstlisting}
  (*$\llbracket$*)@Family interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*) = (*$\llbracket$*)@Obj interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}(*$\rrbracket$*)
\end{lstlisting}
\hspace{.3in}where $\overline{meth'}=\overline{meth}\ \cup\ $\textsf{fieldMethods(}$I_1$\textsf{)}$\ \cup\cdots\cup\ $\textsf{fieldMethods(}$I_n$\textsf{)}, $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)}
~\\~\\
(1) $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)} \textcolor{red}{(definition of newChilds)}
    \begin{itemize}
    \item $\forall I_0\in\overline{I}$, if $\exists I'_0\in\ $\textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} and \textsf{name(}$I_0$\textsf{)} = \textsf{name(}$I'_0$\textsf{)}, then \textsf{combine(}$I_0,I'_0$\textsf{)}$\ \in\overline{I'}$
    \item Otherwise, $I_0\in\overline{I'}$
    \end{itemize}
(2) $\llbracket$\lstinline{@Family}\textsf{ interface }$I$\textsf{ extends }$\overline{I_s},\ \overline{I_t}$\textsf{ \{\}}$\rrbracket\ \in$ \textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} \textcolor{red}{(definition of collectMembers)}
    \begin{itemize}
    \item $\forall i,$ s.t. $I\in$\textsf{ childs(}$I_i$\textsf{)},
        \begin{itemize}
        \item $\overline{I_s}$ = \textsf{suptypes(}$I_i.I$\textsf{)}
        \item $I_i.I\in\overline{I_t}$
        \end{itemize}
    \end{itemize}
(3) $\overline{I_0}$ = \textsf{childs(}$I$\textsf{)}, $\overline{I}$ = \textsf{suptypes(}$I$\textsf{)} \textcolor{red}{(definition of childs and suptypes)}
    \begin{itemize}
    \item \textsf{ibody(}$I$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I}$\textsf{ \{}$\overline{meth}\ \overline{I_0}$\textsf{\}}
    \item \textcolor{red}{need to use ibody and mbody. formalize?}
    \end{itemize}
(4) $meth\in\ $\textsf{fieldMethods(}$I$\textsf{)} \textcolor{red}{(definition of fieldMethods)}
    \begin{itemize}
    \item $meth\in\ $\textsf{childs(}$I$\textsf{)}
    \item $meth$ = $I_0$\textsf{ \_m();}
    \end{itemize}
(5) \textsf{interface }$I$\textsf{ extends }$\overline{I_s}$\textsf{ \{}$\overline{meth}\ \overline{I}$\textsf{\}} = \textsf{combine(}$I_m,I_n$\textsf{)} \textcolor{red}{(definition of combine)}
    \begin{itemize}
    \item \textsf{ibody(}$I_m$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s1}}$\textsf{ \{}$\overline{meth_1}\ \overline{I_1}$\textsf{\}}
    \item \textsf{ibody(}$I_n$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s2}}$\textsf{ \{}$\overline{meth_2}\ \overline{I_2}$\textsf{\}}
    \item $\overline{I_s}$ = $\overline{I_{s1}}\ \cup\ \overline{I_{s2}}$
    \item $\overline{meth}$ = $\overline{meth_1}\ \cup\ \overline{meth_2}$
    \item If $\exists I_1\in\overline{I_1}, I_2\in\overline{I_2}$, \textsf{name(}$I_1$\textsf{)} = \textsf{name(}$I_2$\textsf{)}, then \textsf{combine(}$I_1,I_2$\textsf{)}$\ \in\overline{I}$
    \item Otherwise $(I\in I_1\ \Delta\ I_2)$, $I\in\overline{I}$.
    \end{itemize}
\caption{Translation of \lstinline{@Family}.}
\label{fig:trans1}
\end{figure*}

>>>>

In this section, we present an overview of the code that \lstinline{@Family} generates. The syntax and type system are
consistent with the Java language. We use translation functions to illustrate our code generation.
To make it clearer, we split the process of code generation into three parts, in which case we introduce two
new annotations \lstinline{@Obj} and \lstinline{@ObjOf}, which are not implemented in the source code but just help to explain
the process of translation. In our implementation, the annotation processing is a combination of the three parts, corresponding to
\lstinline{@Family}, \lstinline{@Obj} and \lstinline{@ObjOf} respectively. The translation rules are presented in Figure \ref{fig:trans1}
and Figure \ref{fig:trans2}, together with some auxiliary functions. Note that in practice, our \textsf{@Family} supports simple
generics without bounds, but for simplicity reason it is not included in the formalization.

\begin{figure*}
\begin{tabular}{|l|c|c|}
\hline
\begin{lstlisting}
interface Base {
	interface Exp {}
	interface Lit extends Exp {
		int i();
		Lit withI(int val);
	}
	interface Add extends Exp {
		Exp e1(); Exp e2();
	}
}
\end{lstlisting} & & \\
\hline
\begin{lstlisting}
@Family interface New extends Base {
	interface Exp { int _eval(); }
	interface Lit {
		default int _eval() { return i(); }
	}
	interface Add {
		default int _eval() {
			return e1()._eval() + e2()._eval();
		}
	}
}
\end{lstlisting} & \begin{tabular}{c}$\Rightarrow$\\Family\\Desugaring\end{tabular} & \begin{lstlisting}
@Obj interface New extends Base {
	@Obj interface Exp extends Base.Exp {
		int _eval();
	}
	@Obj interface Lit extends Exp, Base.Lit {
		int i();
		default int _eval() { return i(); }
	}
	@Obj interface Add extends Exp, Base.Add {
		Exp e1(); Exp e2();
		default int _eval() {
			return e1()._eval() + e2()._eval();
		}
	}
}
\end{lstlisting} \\
\hline
& & $\Downarrow$ Obj Desugaring \\
\hline
\begin{lstlisting}
interface New extends Base {
	interface Exp extends Base.Exp {
		int _eval();
	}
	interface Lit extends Exp, Base.Lit {
		int i(); Lit withI(int val);
		default int _eval() { return i(); }
		static Lit of(int _i) {
			return new Lit() {
				int i = _i;
				public int i() { return i; }
				public Lit withI(int val) {
					return Lit.of(val);
				}
			};
		}
	}
	interface Add extends Exp, Base.Add {
		Exp e1(); Exp e2();
		default int _eval() {
			return e1()._eval() + e2()._eval();
		}
		static Add of(Exp _e1, Exp _e2) {
			return new Add() {
				Exp e1 = _e1; Exp e2 = _e2;
				public Exp e1() { return e1; }
				public Exp e2() { return e2; }
			};
		}
	}
	static New of() { return new New(){}; }
}
\end{lstlisting} & \begin{tabular}{c}$\Leftarrow$\\ObjOf\\Desugaring\end{tabular} & \begin{tabular}{l}\begin{lstlisting}
@ObjOf interface New extends Base {
	@ObjOf interface Exp extends Base.Exp {
		int _eval();
	}
	@ObjOf interface Lit extends Exp, Base.Lit {
		int i();
\end{lstlisting}\vspace{-.05in}\\ \begin{lstlisting}[basicstyle=\ttfamily\small\color{red}]
		Lit withI(int val);
\end{lstlisting}\\ \begin{lstlisting}
		default int _eval() { return i(); }
	}
	@ObjOf interface Add extends Exp, Base.Add {
		Exp e1(); Exp e2();
		default int _eval() {
			return e1()._eval() + e2()._eval();
		}
	}
}
\end{lstlisting}\end{tabular} \\
\hline
\end{tabular}
\caption{An example showing the flow of translation. \textcolor{red}{Haoyuan: fields with underscore}}\label{fig:flow}
\end{figure*}

\subsection{Flow of Translation}
Figure~\ref{fig:flow} illustrates the flow of translation with a simple example, showing intermediate results of desugaring. The base family \textsf{Base} contains three members, \textsf{Exp}, and its subtypes \textsf{Lit} and \textsf{Add}. \textsf{Lit} has a field method \textsf{i()} together with a wither method, where \textsf{Add} has two field methods \textsf{e1()} and \textsf{e2()}. At this time, another interface \textsf{New} is defined as the new family that extends the old one, where we add an operation \textsf{\_eval()} for the member types. In the client code, the user simply re-declares the three members with the addition of \textsf{\_eval()}, without building subtyping relations with the old members, and a simple annotation \textsf{@Family} will deal with all the stuff.

In the first step of translation, namely family desugaring, \textsf{@Family} collects all the members from old families, and fills in the ``extends'' to build the inheritance relations. At the same time, it re-declares all the field methods, for example, although the type name of \textsf{e1()} is still \textsf{Exp}, it refers to the new member type, where we address field type refinements, and this new \textsf{e1()} has the operation \textsf{\_eval()}. Then it delegates the work to \textsf{@Obj}. In the second phase, \textsf{@Obj} detects fluent setters and withers from the supertypes, then refines their return types to have consistency with the annotated type. In our example, the wither method \textsf{withI} is refined to have the new \textsf{Lit} as its return type. In the last step of translation, \textsf{@ObjOf} generates constructor methods for all the interfaces, except \textsf{Exp}, where its \textsf{\_eval()} is an abstract method and we cannot provide a default implementation for it. In the following subsections we will respectively explain the three steps of translation in detail.

\subsection{\textsf{@Family}}
\lstinline{@Family} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Family} tackles two
tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
aim, \lstinline{@Family} re-declares all member types and field methods from the base families (see \textsf{collectMembers} and \textsf{fieldMethods} in Figure \ref{fig:trans1}). \lstinline{@Family} recognizes field methods by checking if they are non-void, no-argument methods, and if the name starts with an underscore ``\_''.  The function \textsf{collectMembers} finds all the direct member types from base families $I_1,\cdots,I_n$, creates new types with the same names for re-declaration, and builds the subtyping relations among them. The newly created types are also annotated with \lstinline{@Family}, leading to such a recursive process throughout the nested interfaces in base families. If users have already declared these member types in order to introduce new operations, behaviors or whatever, the function \textsf{combine} will integrate them to the re-declaration, by combining methods and supertypes following ``extends''. Note that we use $\textsf{ibody(}I_m$\textsf{)} to find the exact declaration body of interface $I_m$. \textcolor{red}{Haoyuan: Acutally in the implementation, \textsf{ibody} takes two arguments, one is the name of type reference (interface name $I_m$), and the other is the environment, or where the type reference $I_m$ locates.}

@@

In Figure \ref{fig:trans2}, the first translation function presents the processing of \lstinline{@Obj}. Note that we use $\textsf{mbody(}m,I\textsf{)}$ to get the full declaration of method $m$ seen by $I$, such a method can either be directly defined in $I$ or inherited from its supertypes. Besides delegating \lstinline{@Obj} to the member types in a recursive way, it invokes the function \textsf{refine} to handle the type refinements. The definition of \textsf{refine} is shown in (1) and (2), where return types of \textsf{with-} methods and fluent setters are refined respectively. They are both state operations, and for simplicity we omit the case for general \textsf{with} method in our translation rules.

@@

\begin{figure*}
\begin{lstlisting}
  (*$\llbracket$*)@Family interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*) = (*$\llbracket$*)@Obj interface (*$I_m$*) extends (*$I_1,\cdots,I_n$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}(*$\rrbracket$*)
\end{lstlisting}
\hspace{.3in}where $\overline{meth'}=\overline{meth}\ \cup\ $\textsf{fieldMethods(}$I_1$\textsf{)}$\ \cup\cdots\cup\ $\textsf{fieldMethods(}$I_n$\textsf{)}, $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)}
~\\~\\
(1) $\overline{I'}$ = \textsf{newChilds(}$I_1,\cdots,I_n,\overline{I}$\textsf{)} \textcolor{red}{(definition of newChilds)}
    \begin{itemize}
    \item $\forall I_0\in\overline{I}$, if $\exists I'_0\in\ $\textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} and \textsf{name(}$I_0$\textsf{)} = \textsf{name(}$I'_0$\textsf{)}, then \textsf{combine(}$I_0,I'_0$\textsf{)}$\ \in\overline{I'}$
    \item Otherwise, $I_0\in\overline{I'}$
    \end{itemize}
(2) $\llbracket$\lstinline{@Family}\textsf{ interface }$I$\textsf{ extends }$\overline{I_s},\ \overline{I_t}$\textsf{ \{\}}$\rrbracket\ \in$ \textsf{collectMembers(}$I_1,\cdots,I_n$\textsf{)} \textcolor{red}{(definition of collectMembers)}
    \begin{itemize}
    \item $\forall i,$ s.t. $I\in$\textsf{ childs(}$I_i$\textsf{)},
        \begin{itemize}
        \item $\overline{I_s}$ = \textsf{suptypes(}$I_i.I$\textsf{)}
        \item $I_i.I\in\overline{I_t}$
        \end{itemize}
    \end{itemize}
(3) $\overline{I_0}$ = \textsf{childs(}$I$\textsf{)}, $\overline{I}$ = \textsf{suptypes(}$I$\textsf{)} \textcolor{red}{(definition of childs and suptypes)}
    \begin{itemize}
    \item \textsf{ibody(}$I$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I}$\textsf{ \{}$\overline{meth}\ \overline{I_0}$\textsf{\}}
    \end{itemize}
(4) $meth\in\ $\textsf{fieldMethods(}$I$\textsf{)} \textcolor{red}{(definition of fieldMethods)}
    \begin{itemize}
    \item $meth\in\ $\textsf{childs(}$I$\textsf{)}
    \item $meth$ = $I_0$\textsf{ \_m();}
    \end{itemize}
(5) \textsf{interface }$I$\textsf{ extends }$\overline{I_s}$\textsf{ \{}$\overline{meth}\ \overline{I}$\textsf{\}} = \textsf{combine(}$I_m,I_n$\textsf{)} \textcolor{red}{(definition of combine)}
    \begin{itemize}
    \item \textsf{ibody(}$I_m$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s1}}$\textsf{ \{}$\overline{meth_1}\ \overline{I_1}$\textsf{\}}
    \item \textsf{ibody(}$I_n$\textsf{)} = \textsf{interface }$I$\textsf{ extends }$\overline{I_{s2}}$\textsf{ \{}$\overline{meth_2}\ \overline{I_2}$\textsf{\}}
    \item $\overline{I_s}$ = $\overline{I_{s1}}\ \cup\ \overline{I_{s2}}$
    \item $\overline{meth}$ = $\overline{meth_1}\ \cup\ \overline{meth_2}$
    \item If $\exists I_1\in\overline{I_1}, I_2\in\overline{I_2}$, \textsf{name(}$I_1$\textsf{)} = \textsf{name(}$I_2$\textsf{)}, then \textsf{combine(}$I_1,I_2$\textsf{)}$\ \in\overline{I}$
    \item Otherwise $(I\in I_1\ \Delta\ I_2)$, $I\in\overline{I}$.
    \end{itemize}
\caption{Translation of \lstinline{@Family}.}
\label{fig:trans1}
\end{figure*}
